Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-06-08 ICPP-19/communications.tex
Camera ready version

	\renewcommand{\shortauthors}{Kumar and Eyraud-Dubois, et al.}

	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		%%		\todo[inline]{Initial submission deadline: April 15}
		%With increasing hardware complexity and systems with different memory nodes are ubiquitous in High Performance Computing (HPC). 
		In distributed memory systems, it is paramount to develop strategies to overlap the data transfers between memory nodes with the computations in order to exploit the full potential of these systems. In this paper, we consider the problem of determining the order of data transfers between two memory nodes for a set of independent tasks with the objective of minimizing the makespan. We prove that, with limited memory capacity, the problem of obtaining the optimal order of data transfers is NP-complete. We propose several heuristics to determine this order and discuss the conditions that might be favorable to different heuristics. We analyze our heuristics on traces obtained by running 2 molecular chemistry kernels, namely, Hartree–Fock (HF) and Coupled Cluster Single Double (CCSD) on 10 nodes of an HPC system. Our results show that some of our heuristics achieve significant overlap for moderate memory capacities and resulting in makespans that are very close to the lower bound.
	\end{abstract}

@@

	With the advent of multicore, and the use of accelerators, it is notoriously cumbersome to exploit the full capability of a machine. Indeed, there are several challenges that come into picture. First, every architecture provides its own efficacy and interface. Therefore, a steep learning curve is required for programmers to take good utilization of all resources. Second, scheduling is a well known NP-complete optimization problem, and hybrid and distributed resources make this problem harder (we refer ~\cite{webpagescheduling} for a survey on the complexity of scheduling problems and ~\cite{bleuse2015scheduling} for a recent survey in the case of hybrid nodes). Third, due to shared buses and parallel resources, it is challenging to obtain a precise model based on prediction of computation and communication times. Fourth, the number of architectures has increased drastically in recent years, therefore it is almost impossible to develop hand tuned optimized code for all these architectures. All these observations led to the development of different task based runtime systems. Among several runtimes, we may cite QUARK~\cite{YarKhan:2011:Quark:Manual} and PaRSEC~\cite{parsec} from from ICL, Univ. of Tennessee Knoxville (USA), StarPU~\cite{starpu} form Inria Bordeaux (France), Legion~\cite{legion12} from Stanford Univ. (USA), StarSs~\cite{ompss} from Barcelona Supercomputing Center (Spain), KAAPI~\cite{kaapi} from Inria Grenoble (France). All these runtime systems allow programmers to express their algorithms at the abstract level in the form of direct acyclic graphs (DAG), where vertices represent computations and edges represent dependencies among them. Sometimes some static information is also provided along with the DAG, such as distance to exit (last) node as a priority or affinity of computation towards resources. The runtime is then responsible for managing scheduling of computations and communications, data transfers among different memories, computation-communication overlap, and load balance.

	%%	Historically there is a great focus on the design of parallel algorithms and to minimize the complexity of the computations. Also In the last decade, there is a drastic improvement in the hardware to provide perminent rate of computation, but little improvement has been achieved for the rate of data movement. With extreme scale computing, supercomputers face bottlenecks due to ever growing need of data. Therefore focus of HPC community is now changing towards avoiding, hiding and minimizing communications. (Kyelick work and http://science.energy.gov/~/media/ascr/ ascac/pdf/meetings/20140210/Top10reportFEB14.pdf )

	In the last few decades, we have witnessed a drastic improvement in the hardware to provide higher rate of computation, but comparatively smaller improvement has been achieved for the rate of data movement. With extreme scale computing, supercomputers face bottlenecks due to the need of large amount of data~\cite{ascaccommitteereport2014,yelick2016}. Therefore, the HPC community is now focusing on avoiding, hiding and minimizing communication costs.

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities to computations, based on the task graph structure, and to use them at runtime to make the execution efficient. In irregular applications, programmers do not know the precise structure of the task graphs in advance: tasks are added recursively based on certain sentinel constraints. For such applications, the runtime system sees a set of independent tasks and schedules them on processing units of different memory spaces. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations so as to maximize the overlap between computations and communications. This is the main topic of this paper. We prove that the order of communications on two memory nodes with the objective of minimizing the makespan is a NP-complete problem if the memory of the target node is limited. Our proof is inspired from work by~\cite{Papadimitriou:1980:FSL:322203.322213}, which applies a similar technique for 2-machine flowshop problem with bounded capacity. The main difference between both approaches is that they consider all tasks have the same occupation on the second machine and the memory occupation starts when the processing finishes on the first machine. On the contrary, our approach is designed for tasks appearing in scientific workloads whose memory requirements are highly irregular and we consider that memory is acquired before starting the data transfer on the communication resource. We propose different runtime strategies in order to maximize the overlap of computations and communications. We evaluate our strategies on the context of a cluster of homogeneous nodes. However, our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper:

@@

	The outline of the paper is the following. Section~\ref{sec:relatedWork} describes past work on the computations with limited memory and similar problems in the literature. In section~\ref{sec:theoreticalProof}, we present an algorithm to obtain the order of data transfers when there is not any memory capacity restriction. Then, we also prove that in general data transfer problem is NP-complete. In Section~\ref{sec:heuristics}, we propose several heuristics and discuss the conditions that might be favorable to them. We mainly consider three categories of heuristics: static order based heuristics, dynamic selection based heuristics, and static order with dynamic correction based heuristics. Sections~\ref{sec:expSetting} describes our experimental setup and we evaluate our proposed heuristics on two molecular chemistry kernels in Section~\ref{sec:expResults}. Our results show that static order with dynamic correction based heuristics achieve good performance in most cases. We finally propose conclusions and perspectives in Section~\ref{sec:conclusion}.

@@

	Historically there has been a great emphasis on the development of parallel algorithms and minimizing the complexity of computations. As the number of computation cores has increased drastically in recent years, supercomputers face bottleneck due to communication required by an application. Hence, in recent years the focus has changed towards developing communication avoiding algorithms, strategies to hiding communications and minimizing the data accessed by applications~\cite{yelick2016}.

	The problem of scheduling tasks has been highly studied in the literature and many formulations are known to be NP-complete~\cite{GareyJohnson}. Many static and dynamic strategies have been proposed and analyzed for scheduling tasks on heterogeneous resources~\cite{heft-Topcuoglu,ipdps16starpu}. There is also a number of studies in the direction of task scheduling with the emphasis on improving locality and minimizing the communication cost~\cite{starpu,heft-Topcuoglu}. Stanisic et. al~\cite{luka-dmdar} proposed a heuristic to schedule tasks on a computational resource where most of its data is available. A similar approach has been adopted by Agullo et. al for the scheduling of sparse linear algebra kernels~\cite{agullo_fmm}. Predari et. al proposed heuristics to partition the task graph across a number of processors such that inter-processor communication can be minimized~\cite{predari:tel-01518956}.

	The problem considered in this paper also can be viewed as a flow shop problem: the communication link can be seen as a processing resource, and each task needs to first be handled by the communication link and then by the computational resource. Communication and computation times of a task can thus be considered as processing times on different machines. Johnson has provided scheduling strategies for 2 and 3-machine flow shop problems with infinite memory capacity~\cite{johnson}. 2-machine flow shop problem with finite buffer has been proven NP-complete by Papadimitriou et. al~\cite{Papadimitriou:1980:FSL:322203.322213}, in which a constraint is imposed on the number of tasks that can await execution on the second machine.

	A number of other studies have focused on scheduling with limited memory and storage, starting with the work of register allocation for arithmetic expressions by Sethi and Ulman~\cite{Sethi:1970:GOC:321607.321620}. Sarkar et. al worked on the scheduling of graphs of smaller-grain tasks with limited memory, where each task requires homogeneous data size~\cite{vsarkar-pact}. The same work has been extended by Marchal et. al for task graphs where memory requirement of each task is highly irregular~\cite{loris-ipdps18}.

@@

	To exploit the full potential of a system it may be necessary to execute tasks on processing
	units where all of their data does not reside. A task may require all of its input data
	in local memory before starting the computation. There may be multiple tasks
	scheduled on a processing unit, which require to transfer data from the same
	memory node. Ordering data transfers for such tasks is very crucial for the
	communication-computation overlap, thus for the overall
	performance. In general, order of task execution with input
	and output data transfers can be  
	viewed as a 3-machine flowshop problem, where processing time on the first machine is 
	input data transfer time, processing time on the second machine is task computation time, 
	and processing time on the third machine is output data transfer time; and the objective is 
	to minimize the total makespan. This is a well known NP-complete problem~\cite{3machineFlowShopNPComplete}.

	In many cases, output data that needs to be retrieved after task execution is much smaller than the input data. It is often the case that future tasks running on the same memory node require output data of the past tasks. Therefore, most runtime systems transfer data to other memory nodes based on the demand -- not immediately after they were produced. It is also possible that all output data can be stored in a preallocated separate buffer on a memory node. Hence, we do not consider output data explicitly in our analysis and assume that output data is negligible or stored in a separate buffer for each task. Thus problem considered here is more similar to a 2-machine flowshop problem. We prove that ordering the execution of such tasks with finite memory capacity is a NP-complete problem: %%We also present an optimal algorithm to determine the order of execution of tasks when there is not any restriction on memory limit.

@@

	\vspace*{-0.25cm}
	\begin{algorithm}
		\caption{\label{alg:OrderOfExecutionInfinteMemory}Johnson's~\cite{johnson} algorithm (infinite memory case).}
		\begin{algorithmic}[1]
			\STATE Divide ready tasks in two sets $S_1$ and $S_2$. If computation time of a task $T$ is not less than its communication time, then $T$ is in $S_1$ otherwise in $S_2$.
			\STATE Sort $S_1$ in queue $Q$ by non-decreasing communication times
			\STATE Sort $S_2$ in queue $Q'$ by non-increasing computation times
			\STATE Append $Q'$ to $Q$
			\STATE $\tau_{\text{COMM}} \gets 0$ \hfill\COMMENT{Available time of communication resource}
			\STATE $\tau_{\text{COMP}} \gets 0$\hfill \COMMENT{Available time of computation resource}
			\WHILE{$Q \neq \emptyset$}
			\STATE Remove a task $T$ from beginning of $Q$ for processing
			\STATE $\scomm(T) \gets \tau_{\text{COMM}}$
			\STATE $\scomp(T) \gets max(\scomm(T) + COMM_T, \tau_{\text{COMP}})$
			\STATE $\tau_{\text{COMM}} \gets \scomm(T) + COMM_T$
			\STATE $\tau_{\text{COMP}} \gets \scomp(T) + COMP_T$
			\ENDWHILE
		\end{algorithmic}
	\end{algorithm}
	\vspace*{-0.35cm}
	We also prove optimality of Algorithm~\ref{alg:OrderOfExecutionInfinteMemory} differently in an extended version~\cite{extendedVersion}.
	\subsection{Finite Memory}

	We now consider the general case, in which the memory limit is a
	constraint for the schedule. This is related to previous work by
	Papadimitriou et. al~\cite{Papadimitriou:1980:FSL:322203.322213}, in which the second machine
	can only handle a bounded number of tasks. Our problem generalizes this
	work to heterogeneous memory consumption among tasks, with an
	additional difference: memory usage starts at the beginning of the
	first part of a task (instead of at the end of the first part). This
	requires to provide a slightly different NP-completeness proof, as
	given below.

@@

		In order to prove NP-hardness, we use a reduction from the
		well-known NP-complete problem 3 Partition~\cite{GareyJohnson}:

@@

	This theorem shows that adding a memory constraint to our
	problem makes it more difficult, just like it does for limited capacity 2-machine
	flowshop~\cite{Papadimitriou:1980:FSL:322203.322213}. One additional difficulty of our problem however is that it may not be
	optimal to consider the same order on both machines.

@@

	We classify our heuristics into mainly three categories. In the first category, the order of all computations and communications is computed in advance and the same order is followed on both resources. In the second category, the next task to schedule is dynamically chosen based on different criteria. The final category is based on combining strategies from the first two categories: a static order is precomputed and corrected dynamically to avoid idle time caused by memory limitations. In all of our strategies (except linear programming based strategy), communication and computations take place in the same order.

@@

	We highlight the different dynamic heuristics with the instance described on Table~\ref{tab:dynamicSelectionExample} (these heuristics are too similar on the instance from the previous class), and Figure~\ref{fig:dynamicSelectionExample} shows the corresponding schedules.

@@

	In this class of strategies, we precompute the order of tasks based on some criterion and then follow this ordering as much as possible. But when the communication resource is idle because the memory requirement of the next task is too high, then we select a task  with a dynamic strategy. After a task is dynamically selected, we update the remaining order without this task. This class of strategies takes advantage of static information in the form of precomputed order and dynamic correction to minimize the idle time due to memory constraint.

@@

	%%\begin{figure}[htb]
	%%\includegraphics[scale=0.05]{Figs/staticOrderDynamiccorrectionSchedules}
	%%\caption{Task set and schedules where orders are precomputed based on different heuristics.}
	%%\label{fig:staticOrderDynamiccorrectionExample}
	%%\end{figure}
\begin{figure}[!htb]
	\newcommand{\taskA}[2][0]{\task[#1]{$A$}{#2}{4}{1}{cyan}}
	\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{2}{6}{cyan!50!black}}
	\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{8}{8}{blue!40!white}}
	\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{5}{4}{blue!70!white}}
	\newcommand{\taskE}[2][0]{\task[#1]{$E$}{#2}{3}{2}{blue}}
	\vspace*{-0.35cm}
	\begin{tikzpicture}[yscale=0.5,xscale=0.2]
	\begin{scope}
	\schedule{35}{OOLCMR}{2,8,12,15,17,25,33}
	\taskB{0}
	\taskD[1]{2}
	\taskA{8}
	\taskE{12}
	\taskC{17}
	\end{scope}
	\begin{scope}[yshift=-2.75cm]
	%%		\schedule{36}{OOSCMR}{2,5,8,11,14,18,26,34}
	%%		\taskB{0}
	%%		\taskE[3]{2}
	%%		\taskA[1]{5}
	%%		\taskD{9}
	%%		\taskC{18}
	%%		\end{scope}
	\schedule{36}{OOSCMR}{2,5,8,11,15,19,27,35}
	\taskB{0}
	\taskE[3]{2}
	\taskA[1]{5}
	\taskD{10}
	\taskC{19}
	\end{scope}
	\begin{scope}[yshift=-5.5cm]
	\schedule{36}{OOMAMR}{2,8,11,14,17,25,33}
	\taskB{0}
	\taskD[1]{2}
	\taskE[1]{8}
	\taskA{12}
	\taskC{17}
	\end{scope}
	%%		\begin{scope}[yshift=+2.75cm]
	%%		\schedule{40}{OMIM}{2,8,16,24,29,33,38}
	%%		\taskB{0}
	%%		\taskC{8}
	%%		\taskD{24}
	%%		\taskA{29}
	%%		\taskE{33}
	%%		\end{scope}

	\end{tikzpicture}
	\caption{ \label{fig:staticOrderDynamiccorrectionExample} Different static order dynamic correction heuristic schedules for a task set of Table~\ref{tab:staticOrderDynamiccorrectionExample} with a memory capacity of 9. The $OMIM$ order is $BCDEA$.\vspace*{-0.5cm}}
\end{figure}

\begin{table}[!htb]
	\begin{center}
		\begin{tabular}{|c|c|c|c|}
			\hline
			Task & Memory Req & Comm Time & Comp Time\\ \hline 
%%			\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
%%			&=Comm Volume && \\ \hline
			A & 4 & 4 &  1\\ \hline
			B & 2 & 2 & 6\\ \hline
			C & 8 & 8 & 8\\ \hline
			D & 5 & 5 & 4\\ \hline
			E & 3 & 3 & 2\\ \hline
		\end{tabular}
		\caption{\label{tab:staticOrderDynamiccorrectionExample} A task set for static order dynamic correction schedules.\vspace*{-0.75cm}}
	\end{center}
\end{table}

	As previously, we propose on Table~\ref{tab:staticOrderDynamiccorrectionExample} an example instance for this class of heuristics, and provide on Figure~\ref{fig:staticOrderDynamiccorrectionExample} the corresponding schedules.

@@

We also consider two other static heuristics for evaluation. The first heuristic is based on an algorithm, proposed by Gilmore and Gomory, to obtain the minimal cost sequence for a set of jobs~\cite{Gilmore-Gomory:1964}. This is a classical algorithm for 2-machine no-wait flow shop problem. In this algorithm, each job has a start and end state and a cost is associated to change the state. In our context, this cost can be seen as non-overlap time of computation for two adjacent tasks. Here is the main idea of this algorithm. Initially, a partial sequence of jobs is represented by a graph such that their overlap is maximum. Subsequently edges are greedily added to this graph to connect two components while minimizing the total non-overlap cost. When the graph is  connected, an edge interchange mechanism is used to determine the sequence of jobs, which ensures that the sequence has minimal cost. More details can be found in the original paper~\cite{Gilmore-Gomory:1964} and our implementation is publicly available~\cite{gitworkrepo}. This algorithm does not take memory constraints into account and only provides the sequence of processing. We use this sequence with a memory capacity restriction just like for other static heuristics, and we call this heuristic \textit{Gilmore-Gomory} ($GG$).

@@

		\section{Conclusion and Perspectives}
		\label{sec:conclusion}

>>>>

	\renewcommand{\shortauthors}{Kumar, Eyraud-Dubois, and Krishnamoorthy}

	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		%%		\todo[inline]{Initial submission deadline: April 15}
		%With increasing hardware complexity and systems with different memory nodes are ubiquitous in High Performance Computing (HPC). 
		In distributed memory systems, it is paramount to develop strategies to overlap the data transfers between memory nodes with the computations in order to exploit their full potential. In this paper, we consider the problem of determining the order of data transfers between two memory nodes for a set of independent tasks with the objective of minimizing the makespan. We prove that, with limited memory capacity, the problem of obtaining the optimal data transfer order is NP-complete. We propose several heuristics to determine this order and discuss the conditions that might be favorable to different heuristics. We analyze our heuristics on traces obtained by running two molecular chemistry kernels, namely, Hartree–Fock (HF) and Coupled Cluster Singles Doubles (CCSD), on 10 nodes of an HPC system. Our results show that some of our heuristics achieve significant overlap for moderate memory capacities and resulting in makespans that are very close to the lower bound.
	\end{abstract}

@@

	With the advent of multicore, and the use of accelerators, it is notoriously cumbersome to exploit the full capability of a machine. Indeed, there are several challenges that come into picture. First, every architecture provides its own efficacy and interface. Therefore, a steep learning curve is required for programmers to effectively utilize all resources. Second, scheduling is a well known NP-complete optimization problem, and hybrid and distributed resources make this problem harder (we refer \citet{webpagescheduling} for a survey on the complexity of scheduling problems and \citet{bleuse2015scheduling} for a recent survey in the case of hybrid nodes). Third, due to shared buses and parallel resources, it is challenging to obtain a precise model based on prediction of computation and communication times. Fourth, the number of architectures has increased drastically in recent years, making it almost impossible to develop hand-tuned optimized code for all these architectures. All these observations led to the development of different task-based runtime systems. Among several runtimes, we may cite QUARK~\cite{quark} and PaRSEC~\cite{parsec} from  ICL, Univ. of Tennessee Knoxville (USA), StarPU~\cite{starpu} from Inria Bordeaux (France), Legion~\cite{legion12} from Stanford Univ. (USA), StarSs~\cite{ompss} from Barcelona Supercomputing Center (Spain), KAAPI~\cite{kaapi} from Inria Grenoble (France). All these runtime systems allow programmers to express their algorithms at the abstract level in the form of direct acyclic graphs (DAG), where vertices represent computations and edges represent dependencies among them. Sometimes some static information is also provided along with the DAG, such as distance to exit (last) node as a priority or affinity of computation toward resources. The runtime is then responsible for managing the schedule of computations and communications, data transfer among different memories, computation-communication overlap, and load balance.

	%%	Historically there is a great focus on the design of parallel algorithms and to minimize the complexity of the computations. Also In the last decade, there is a drastic improvement in the hardware to provide perminent rate of computation, but little improvement has been achieved for the rate of data movement. With extreme scale computing, supercomputers face bottlenecks due to ever growing need of data. Therefore focus of HPC community is now changing toward avoiding, hiding and minimizing communications. (Kyelick work and http://science.energy.gov/~/media/ascr/ ascac/pdf/meetings/20140210/Top10reportFEB14.pdf )

	In the last few decades, we have witnessed a drastic improvement in the hardware to provide higher rate of computation, but comparatively smaller improvement has been achieved for the rate of data movement. With extreme-scale computing, supercomputers face bottlenecks due to the need for large amount of data~\cite{ascaccommitteereport2014,yelick2016}. Therefore, the HPC community is now focused on avoiding, hiding, and minimizing communication costs.

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities to computations, based on the task graph structure, and to use them at runtime to make the execution efficient. In irregular applications, programmers do not know the precise structure of the task graphs in advance: tasks are added recursively based on certain sentinel constraints. For such applications, the runtime system sees a set of tasks and schedules them on processing units of different memory spaces. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations so as to maximize the overlap between computation and communication. This is the main topic of this paper. We prove that determining the order of communications on two memory nodes with the objective of minimizing the makespan is a NP-complete problem if the memory of the target node is limited. Our proof is inspired by~\cite{Papadimitriou:1980:FSL:322203.322213}, which applies a similar technique for 2-machine flowshop problem with bounded capacity. The main difference between both approaches is that they consider all tasks have the same occupation on the second machine and the memory occupation starts when the processing finishes on the first machine. On the contrary, our approach is designed for tasks appearing in scientific workloads whose memory requirements are highly irregular and we consider that memory is acquired before starting the data transfer on the communication resource. We propose different runtime strategies in order to maximize the  computation-communication overlap. We evaluate our strategies on the context of a cluster of homogeneous nodes. However, our approach is generic and easily adaptable to any system that operates on different memory spaces. Here are the important contributions of this paper:

@@

	The rest of the paper is organized as follows. Section~\ref{sec:relatedWork} describes past work on the computations with limited memory and similar problems in the literature. In Section~\ref{sec:theoreticalProof}, we present an algorithm to obtain the order of data transfers when there is not any memory capacity restriction. Also, we prove that, in general, the data transfer problem is NP-complete. In Section~\ref{sec:heuristics}, we propose several heuristics and discuss the conditions that might be favorable to them. We mainly consider three categories of heuristics: static order based heuristics, dynamic selection based heuristics, and static order with dynamic correction based heuristics. Sections~\ref{sec:expSetting} describes our experimental setup and we evaluate our proposed heuristics on two molecular chemistry kernels in Section~\ref{sec:expResults}. Our results show that static order with dynamic correction based heuristics achieve good performance in most cases. We finally present conclusions and perspectives in Section~\ref{sec:conclusion}.

@@

	Historically there has been a great emphasis on the development of parallel algorithms and minimizing the complexity of computations. As the number of computation cores has increased in recent years, supercomputers face bottlenecks due to the communication required by an application. Hence, the focus has shifted toward developing communication avoiding algorithms, strategies to hiding communications, and minimizing the data accessed by applications~\cite{yelick2016}.

	The problem of scheduling tasks has been highly studied in the literature and many formulations are known to be NP-complete~\cite{GareyJohnson}. Many static and dynamic strategies have been proposed and analyzed for scheduling tasks on heterogeneous resources~\cite{heft-Topcuoglu,ipdps16starpu, hipc16multiresource}. There is also a number of studies in the direction of task scheduling with the emphasis on improving locality and minimizing the communication cost~\cite{starpu,heft-Topcuoglu}. Stanisic et al.~\cite{luka-dmdar} proposed a heuristic to schedule tasks on a computational resource where most of its data is available. A similar approach has been adopted by Agullo et al. for the scheduling of sparse linear algebra kernels~\cite{agullo_fmm}. Predari et al. proposed heuristics to partition the task graph across a number of processors such that inter-processor communication can be minimized~\cite{predari:tel-01518956}.

	The problem we consider also can be viewed as a flowshop problem: the communication link can be seen as a processing resource, and each task needs to first be handled by the communication link and then by the computational resource. Communication and computation times of a task can thus be considered as processing times on different machines. Johnson presented scheduling strategies for 2 and 3-machine flowshop problems with infinite memory capacity~\cite{johnson}. Two-machine flowshop problem with finite buffer was proven NP-complete by %Papadimitriou et
	\citet{Papadimitriou:1980:FSL:322203.322213}, in which a constraint is imposed on the number of tasks that can await execution on the second machine.

	A number of other studies have focused on scheduling with limited memory and storage, starting with the work of register allocation for arithmetic expressions by Sethi and Ulman~\cite{Sethi:1970:GOC:321607.321620}. Sarkar et al. worked on the scheduling of graphs of smaller-grain tasks with limited memory, where each task requires homogeneous data size~\cite{vsarkar-pact}. The same work has been extended by Marchal et al. for task graphs where per-task memory requirements are highly irregular~\cite{loris-ipdps18}.

@@

	To exploit the full potential of a system it may be necessary to execute tasks on processing
	units where all of their data does not reside. A task may require all of its input data
	in local memory before starting the computation. There may be multiple tasks
	scheduled on a processing unit, which require to transfer data from the same
	memory node. Ordering data transfers for such tasks is very crucial for the
	communication-computation overlap, thus for the overall
	performance. In general, order of task execution with input
	and output data transfers can be  
	viewed as a 3-machine flowshop problem, where processing time on the first machine is 
	input data transfer time, processing time on the second machine is task computation time, 
	and processing time on the third machine is output data transfer time; and the objective is 
	to minimize the total makespan. This is a well-known NP-complete problem~\cite{3machineFlowShopNPComplete}.

	In many cases, output data to be retrieved after task execution is much smaller than the input data. It is often the case that future tasks running on the same memory node require output data of the past tasks. Therefore, most runtime systems transfer data to other memory nodes based on the demand -- not immediately after they were produced. It is also possible that all output data can be stored in a preallocated separate buffer on a memory node. We do not consider output data explicitly in our analysis and assume that output data is negligible or stored in a separate buffer for each task. Thus the problem we consider is more similar to a 2-machine flowshop problem. We prove that ordering the execution of such tasks with finite memory capacity is a NP-complete problem: %%We also present an optimal algorithm to determine the order of execution of tasks when there is not any restriction on memory limit.

@@

	\vspace*{-0.25cm}
	\begin{algorithm}
		\caption{\label{alg:OrderOfExecutionInfinteMemory}Johnson's~\cite{johnson} algorithm (infinite memory case).}
		\begin{algorithmic}[1]
			\STATE Divide ready tasks in two sets $S_1$ and $S_2$. If computation time of a task $T$ is not less than its communication time, then $T$ is in $S_1$ otherwise in $S_2$.
			\STATE Sort $S_1$ in queue $Q$ by non-decreasing communication times
			\STATE Sort $S_2$ in queue $Q'$ by non-increasing computation times
			\STATE Append $Q'$ to $Q$
			\STATE $\tau_{\text{COMM}} \gets 0$ \hfill\COMMENT{Available time of communication resource}
			\STATE $\tau_{\text{COMP}} \gets 0$\hfill \COMMENT{Available time of computation resource}
			\WHILE{$Q \neq \emptyset$}
			\STATE Remove a task $T$ from beginning of $Q$ for processing
			\STATE $\scomm(T) \gets \tau_{\text{COMM}}$
			\STATE $\scomp(T) \gets max(\scomm(T) + COMM_T, \tau_{\text{COMP}})$
			\STATE $\tau_{\text{COMM}} \gets \scomm(T) + COMM_T$
			\STATE $\tau_{\text{COMP}} \gets \scomp(T) + COMP_T$
			\ENDWHILE
		\end{algorithmic}
	\end{algorithm}
	\vspace*{-0.35cm}
	We provide an alternative optimality proof of Algorithm~\ref{alg:OrderOfExecutionInfinteMemory} in the extended version~\cite{extendedVersion}.
	\subsection{Finite Memory}

	We now consider the general case, in which the memory limit is a
	constraint for the schedule. This is related to previous work by
	%Papadimitriou et al.~ 
	\citet{Papadimitriou:1980:FSL:322203.322213}, in which the second machine
	can only handle a bounded number of tasks. Our problem generalizes this
	work to heterogeneous memory consumption among tasks, with an
	additional difference: memory usage starts at the beginning of the
	first part of a task (instead of at the end of the first part). This
	requires us to provide a slightly different NP-completeness proof, as
	given below.

@@

		To prove NP-hardness, we use a reduction from the
		well-known 3-Partition NP-complete problem ~\cite{GareyJohnson}:

@@

	This theorem shows that adding a memory constraint to our
	problem makes it more difficult, just like it does for limited capacity 2-machine
	flowshop~\cite{Papadimitriou:1980:FSL:322203.322213}. One additional difficulty for our problem is that it may not be
	optimal to consider the same order on both machines.

@@

	We classify our heuristics into mainly three categories. In the first category, the order of all computations and communications is computed in advance and the same order is followed on both resources. In the second category, the next task to schedule is dynamically chosen based on different criteria. The final category is based on combining strategies from the first two categories: a static order is precomputed and corrected dynamically to avoid idle time caused by memory limitations. In all of our strategies (except linear programming based strategy), communication and computation orders of tasks are same.

@@

	Similar to the schedules of the previous class, Figure~\ref{fig:dynamicSelectionExample}  highlights the  schedules of this class of heuristics for the example instance of Table~\ref{tab:dynamicSelectionExample}.
	%%	We highlight the different dynamic heuristics with the instance described on Table~\ref{tab:dynamicSelectionExample} (these heuristics are similar on the instance from the previous class), and Figure~\ref{fig:dynamicSelectionExample} shows the corresponding schedules.

@@

	In this class of strategies, we precompute the order of tasks based on some criterion and then follow this ordering as much as possible. But when the communication resource is idle because the memory requirement of the next task is too high, then we select a task  with a dynamic strategy. After a task is selected, we update the remaining order without this task. This class of strategies takes advantage of static information in the form of precomputed order and dynamic correction to minimize the idle time due to memory constraint.

@@

%%	\vspace*{0.05cm}
	%%\begin{figure}[htb]
	%%\includegraphics[scale=0.05]{Figs/staticOrderDynamiccorrectionSchedules}
	%%\caption{Task set and schedules where orders are precomputed based on different heuristics.}
	%%\label{fig:staticOrderDynamiccorrectionExample}
	%%\end{figure}
	\begin{figure}[!htb]
		\newcommand{\taskA}[2][0]{\task[#1]{$A$}{#2}{4}{1}{cyan}}
		\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{2}{6}{cyan!50!black}}
		\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{8}{8}{blue!40!white}}
		\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{5}{4}{blue!70!white}}
		\newcommand{\taskE}[2][0]{\task[#1]{$E$}{#2}{3}{2}{blue}}
		\vspace*{-0.35cm}
		\begin{tikzpicture}[yscale=0.5,xscale=0.2]
		\begin{scope}
		\schedule{35}{OOLCMR}{2,8,12,15,17,25,33}
		\taskB{0}
		\taskD[1]{2}
		\taskA{8}
		\taskE{12}
		\taskC{17}
		\end{scope}
		\begin{scope}[yshift=-2.75cm]
		%%		\schedule{36}{OOSCMR}{2,5,8,11,14,18,26,34}
		%%		\taskB{0}
		%%		\taskE[3]{2}
		%%		\taskA[1]{5}
		%%		\taskD{9}
		%%		\taskC{18}
		%%		\end{scope}
		\schedule{36}{OOSCMR}{2,5,8,11,15,19,27,35}
		\taskB{0}
		\taskE[3]{2}
		\taskA[1]{5}
		\taskD{10}
		\taskC{19}
		\end{scope}
		\begin{scope}[yshift=-5.5cm]
		\schedule{36}{OOMAMR}{2,8,11,14,17,25,33}
		\taskB{0}
		\taskD[1]{2}
		\taskE[1]{8}
		\taskA{12}
		\taskC{17}
		\end{scope}
		%%		\begin{scope}[yshift=+2.75cm]
		%%		\schedule{40}{OMIM}{2,8,16,24,29,33,38}
		%%		\taskB{0}
		%%		\taskC{8}
		%%		\taskD{24}
		%%		\taskA{29}
		%%		\taskE{33}
		%%		\end{scope}

		\end{tikzpicture}
		\vspace*{-0.25cm}\caption{ \label{fig:staticOrderDynamiccorrectionExample} Different static order dynamic correction heuristic schedules for a task set of Table~\ref{tab:staticOrderDynamiccorrectionExample} with a memory capacity of 9. The $OMIM$ order is $BCDEA$.\vspace*{-0.25cm}}
	\end{figure}
%%\vspace*{2mm}
	\begin{table}[!htb]
%%		\begin{center}
\vspace*{-0.25cm}
			\begin{tabular}{|c|c|c|c|}
				\hline
				Task & Memory Req & Comm Time & Comp Time\\ \hline 
				%%			\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
				%%			&=Comm Volume && \\ \hline
				A & 4 & 4 &  1\\ \hline
				B & 2 & 2 & 6\\ \hline
				C & 8 & 8 & 8\\ \hline
				D & 5 & 5 & 4\\ \hline
				E & 3 & 3 & 2\\ \hline
			\end{tabular}\vspace*{-0.05cm}
			\caption{\label{tab:staticOrderDynamiccorrectionExample}A task set for static order dynamic correction schedules.\vspace*{-0.85cm}}
%%		\end{center}
	\end{table}	
	%%	As previously, we propose on Table~\ref{tab:staticOrderDynamiccorrectionExample} an example instance for this class of heuristics, and provide on Figure~\ref{fig:staticOrderDynamiccorrectionExample} the corresponding schedules.
	As previously, we propose on Table~\ref{tab:staticOrderDynamiccorrectionExample} an example instance for this class of heuristics, and the corresponding schedules in Figure~\ref{fig:staticOrderDynamiccorrectionExample}.

@@

	We also consider two other static heuristics for evaluation. The first heuristic is based on an algorithm, proposed by Gilmore and Gomory, to obtain the minimal cost sequence for a set of jobs~\cite{Gilmore-Gomory:1964}. This is a classical algorithm for 2-machine no-wait flowshop problem. In this algorithm, each job has a start and end state and a cost is associated to change the state. In our context, this cost can be seen as non-overlap time of computation for two adjacent tasks. Here is the main idea of this algorithm. Initially, a partial sequence of jobs is represented by a graph such that their overlap is maximum. Subsequently edges are greedily added to this graph to connect two components while minimizing the total non-overlap cost. When the graph is  connected, an edge interchange mechanism is used to \newline determine the sequence of jobs, which ensures that the sequence has minimal cost. We refer the original paper~\cite{Gilmore-Gomory:1964} for more details.
	%% and our implementation is publicly available~\cite{gitworkrepo}. 
	This algorithm does not take memory constraints into account and only provides the sequence of processing. We use this sequence with a memory capacity restriction just like for other static heuristics, and we call this heuristic \textit{Gilmore-Gomory} ($GG$).

@@

		\section{Conclusions and Perspectives}
		\label{sec:conclusion}
